Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Guideline for International \emph{and Domestic} Students

### 2024-02-21 demystify.tex
updates

Most PhD programs will require at least \textbf{two LORs}. Having a letter from an internationally recognized researcher will greatly strengthen your application. In fact, a strong letter from a well-known researcher can outweigh the lack of publications.

@@

\paragraph{Self-written Letters} Many international students write letters themselves, typically due to the requests of their profs., and have them signed by their profs. Such letters have little values and are considered weak by reviewers (why can you not even find someone who cares or knows enough about you to write a candid personal reference letter?).  Instead of the reference writer talking about you, in this case it is you who write about yourself (and they just sign the letter).

@@

\paragraph{Interviews} Many PhD programs have the tradition of interviewing applicants (just like job applications). The most common case is that a professor is interested in working with a student and want to chat with them to make decision (e.g., to offer them RA).
The professor might include one of their current students as part of the interview.

In some case, the interview is done by several professors, e.g., to see if a student fits to their group or to recruit a very strong student to their program. The interview can also be done by Adcom members to see if the student fit their program.

@@

\begin{commentbox}[Vu:]
  At GMU, faculty are encouraged to interview candidates. For very strong candidates, the interview is actually to recruit them.  In some cases a faculty interviews a candidate that they see potentials and want to advocate for their admission. Without the interview, such applications may be more likely to be rejected.
  \tcblower
  In short, getting an interview is a good sign; it means that someone is considering you. If we are not interested in your application, we will not proceed with an interview.

@@

        You can then share this sheet with your reference writers and remind them to submit LoRs on your behalf (see \autoref{sec:lor}).
\end{itemize}

@@

By around March you should hear back from most PhD programs you applied to. If you haven't heard back, reach out via email and ask about the status of your application.
If you receive offers, congratulations!  Now you're at a different game because the schools that have admitted you will try to get you to accept them!  Important factors to consider are the reputation of schools and professors (\autoref{sec:schoolsandprofs}), and funding availability (\autoref{sec:funding}). You will have to make your decision by around \emph{April 15}.

@@

>>>>

Most PhD programs will require at least \textbf{two LORs}. Having a strong letter from an internationally recognized researcher will greatly strengthen your application. In fact, a strong letter from a well-known researcher can outweight the lack of publications.

@@

\paragraph{Self-written Letters} Many international students write letters themselves, typically due to the requests of their profs., and have them signed by their profs. Such letters have little values and are considered weak by reviewers---why can you not even find someone who cares or knows enough about you to write a candid personal reference letter?.  Instead of the reference writer talking about you, in this case it is you who write about yourself (and they just sign the letter).

@@

\paragraph{Interviews} Just like a job application, you can get interviews from some schools. The most common case is that a professor is interested in working with you and want to chat to make decision (e.g., to offer RA).
The professor might include one of their current students as part of the interview.

In some case, the interview is done by several professors, e.g., to see if a student fits to their group or to recruit a very strong student to their program. The interview can also be done by adcom members to see if the student fit their program.

@@

\begin{commentbox}[Vu:]
  At GMU, faculty are encouraged to interview candidates. For very strong candidates, the interview is actually to recruit them.  In some cases a faculty interviews a candidate that they see potentials and want to advocate for their admission. Without the interview, such applications may be more likely to be rejected.
  \tcblower
  In short, getting an interview is a good sign; it means that someone is considering you. If we are not interested in your application, we will not proceed with an interview.
\end{commentbox}

@@

\end{itemize}

You can then share this sheet with your reference writers and remind them to submit LoRs on your behalf (see \autoref{sec:lor}). Also, you should periodically update your writers with your status; a simple note such as \emph{``I am getting interviewed by prof. X at school Y. Do you have any advice?''} can get you a lot of help.

@@

By around March you should hear back from most PhD programs you applied to. If you haven't heard back, reach out via email and ask about the status of your application.
If you receive offers, congratulations!  Now you're at a different game because the schools that have admitted you will try to get you to accept them!  Important factors to consider include the reputation of schools and professors (\autoref{sec:schoolsandprofs}), and funding availability (\autoref{sec:funding}). You will have to make your decision by around \emph{April 15}.

@@

\section{Buying Computer Equipments}\label{sec:buying-equipment} Many students, including myself, get very excited about getting to buy new computer equipments and electronics when they start their PhD. However, you should check with your adviser first.  Many profs. have funding to buy you a computer and other equipments (e.g., keyboards, headphones, tablets, etc).  Also, do not worry that this is a "waste" of your adviser's money.  This is part of the funding that your adviser gets to support you and your research.  So use it to get what you need to do your research (and it's also OK to use them for personal purposes!).
